Niche model food webs and stability
========================================================

Load required libraries
```{r}
require(devtools)
require(igraph)
require(NetIndices)
```

Use the niche model function created by Ted Hart of Distributed Ecology
```{r}
niche.model<-function(S,C){
  new.mat<-matrix(0,nrow=S,ncol=S)
	ci<-vector()
	niche<-runif(S,0,1)
	r<-rbeta(S,1,((1/(2*C))-1))*niche
	
	for(i in 1:S){
		ci[i]<-runif(1,r[i]/2,niche[i])
	}
	
	r[which(niche==min(niche))]<-.00000001
	
	for(i in 1:S){
		
		for(j in 1:S){
			if(niche[j]>(ci[i]-(.5*r[i])) && niche[j]<(ci[i]+.5*r[i])){
				new.mat[j,i]<-1
			}
		}
	}

	new.mat<-new.mat[,order(apply(new.mat,2,sum))]
	return(new.mat)
}

```

Set the desired levels of number of nodes and connectances
```{r}
nodes <- seq(10, 100, 10)
connectance <- seq(.05, .45, .025)
```

Create a data frame of node/connectance pairings
```{r}
niche.inputs <- data.frame(N = rep(nodes, each = 17), C = rep(connectance, 10))
```

Generate 10 niche model food webs for each pairing of node and connectance
```{r}
web_maker <- function(n, x){ # where n is # of webs to make for each pairing and x is a data frame of node/connectance pairings
  # Make sure the column names in the data frame are N and C 
  if(!colnames(x[1]) == "N"){
    warning("Check column names")
  }
  if(!colnames(x[2]) == "C"){
    warning("Check column names")
  }
  
  web <- list()
  
  for(i in 1:length(x[,1])){
    subweb <- list()
    for(j in 1:n){
      subweb[[j]] <- niche.model(x$N[i], x$C[i])
    }
    web[[i]] <- subweb
  }
  
  return(web)
}
```

Create a list of niche model food webs for each pair of parameters
```{r}
system.time(
  test1 <- web_maker(10, niche.inputs)
  )
```

Call eigenvalue stability analysis functions from GitHub repository
```{r}
source("~/Desktop/GitHub/Ecological-Networks/ChainLength/Rscripts/chains_functions.R")
```

Define the parameters for `find_qss`
```{r}
params.u <- data.frame(pred1 = 0, pred2 = 10, prey1 = 0, prey2 = 1)
```

Analyze the quasi-sign stability of the niche model food webs with `find_qss`
```{r}
system.time(
  test2 <- sapply(test1, find_qss, mode = "unif", parms = params.u, iter = 1000)
)
colnames(test2) <- paste(niche.inputs$N, niche.inputs$C)
```